---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- used devtools::build_readme() to update the md -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# afrilearndata

<!-- badges: start -->
<!-- badges: end -->

afrilearndata provides small African datasets to help with learning and teaching of spatial techniques and mapping.

## Installation

Install the development version of afrilearndata with:

``` r

    # install.packages("remotes") # if not already installed
    
    remotes::install_github("afrimapr/afrilearndata")
    
```

## First use

The package contains the following objects 

1. `sfafricountries`    polygons, country boundaries
1. `sfafricontinent`    polygons, continent outline including madagascar
1. `sfafrihway`         lines, trans african highway network
1. `sfafricapitals`     points, capital cities
1. `rastafriwpop`       raster grid, modelled population density per square km aggregated to 20km squares

Lazy loading means that the objects should be accessible once `library(afrilearndata)` is used.

If they are not recognised you can use e.g. `data(sfafricountries)` to make sure the objects are loaded.

```{r countries}

library(afrilearndata)
library(sf)

# polygons
plot(sf::st_geometry(sfafricountries))

```

```{r highway}

# lines
plot(sf::st_geometry(sfafrihway))

```

```{r capitals}

# points
plot(sf::st_geometry(sfafricapitals))

```


```{r population grid}

# raster grid
# install.packages("raster") # if not already installed
library(raster)
plot(rastafriwpop)

```

Interactive maps can be created using the `mapview` package.

```{r interactive}

# install.packages("mapview") # if not already installed
if (FALSE) {
  
  library(mapview)
  mapview::mapview(sfafricountries, zcol="name")  
  
}


```


