---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- used devtools::build_readme() to update the md -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# afrilearndata

<!-- badges: start -->
<!-- badges: end -->

afrilearndata provides small African datasets to help with learning and teaching of spatial techniques and mapping.

## Installation

Install the development version of afrilearndata with:

```{r setup, eval=FALSE, message=FALSE, warning=FALSE}

    # install.packages("remotes") # if not already installed
    
    remotes::install_github("afrimapr/afrilearndata")
    
    library(afrilearndata)

```

## Outline

The package contains the following objects 

1. `africountries`    polygons, country boundaries
1. `africontinent`    polygons, continent outline including madagascar
1. `afrihighway`         lines, trans african highway network
1. `africapitals`     points, capital cities
1. `afriairports`     points, airports
1. `afripop2020`   raster grid, population density 2020 from [WorldPop](https://www.worldpop.org/) aggregated to 20km squares
1. `afripop2000`   raster grid, population density 2000 from [WorldPop](https://www.worldpop.org/) aggregated to 20km squares

Lazy loading means that the objects should be accessible once `library(afrilearndata)` is used.

If they are not recognised you can use e.g. `data(africountries)` to make sure the objects are loaded.


Firstly, here are most of the data shown together. The `tmap` code to create this plot is shown later in the readme. 

```{r tmap-all-the-data, echo=FALSE, message=FALSE, warning=FALSE}
#echo=FALSE to not show code in first version at top of readme
#do show the code later

library(afrilearndata)

# install.packages("tmap") # if not already installed
library(tmap)

# tmap_mode("view") to set to tmap interactive viewing mode

tm_shape(afripop2020) +
    tm_raster(palette = rev(viridisLite::magma(5)), breaks=c(0,2,20,200,2000,25000)) +
tm_shape(africountries) +
    tm_borders("white", lwd = .5) +
    #tm_text("iso_a3", size = "AREA") +
tm_shape(afrihighway) +
    tm_lines(col = "black") + 
tm_shape(africapitals) +
    tm_symbols(col = "blue", alpha=0.4, scale = .6 )+
tm_legend(show = FALSE)

```

Now looking at the data layers individually plotted with packages `sf` or `raster`

```{r countries, message=FALSE, warning=FALSE}

library(afrilearndata)
library(sf)

# polygons
plot(sf::st_geometry(africountries))

```

```{r highway, message=FALSE, warning=FALSE}

# lines
plot(sf::st_geometry(afrihighway))

```

```{r capitals, message=FALSE, warning=FALSE}

# points
plot(sf::st_geometry(africapitals))

```


Population density data are from WorldPop clipped to Africa and aggregated to 20km resolution to make them more manageable. [WorldPop](https://www.worldpop.org/) datasets are licensed under [Creative Commons Attribution 4.0 International](https://creativecommons.org/licenses/by/4.0/).


```{r population-grid, message=FALSE, warning=FALSE}

# raster grid
# install.packages("raster") # if not already installed
library(raster)
plot(afripop2020)

```

The `africountries` data has country names in French, Portuguese, Swahili, Afrikaans and English, that can be used to label maps as follows.

```{r french-country-names, warnings=FALSE, eval=TRUE, echo=TRUE}

library(afrilearndata)

# install.packages("tmap") # if not already installed
library(tmap)

tm_shape(africountries) +
     tm_borders("grey", lwd = .5) +
     tm_text("name_fr", auto.placement=FALSE, remove.overlap=FALSE, just='centre', col='red4', size=0.7 )
```

Interactive maps can be created using the `mapview` package.

```{r interactive, eval=FALSE}

# install.packages("mapview") # if not already installed

library(mapview)
mapview::mapview(africountries, zcol="name")
#here to show all airports on the continent
mapview(afriairports, zcol='type', label='name', cex=2)
  
```

Here is a repeat of the map shown at the start of the readme, together with the code used to create it.

```{r tmap-code, warnings=FALSE, eval=TRUE, echo=TRUE}

library(afrilearndata)

# install.packages("tmap") # if not already installed
library(tmap)

# tmap_mode("view") to set to tmap interactive viewing mode

tm_shape(afripop2020) +
    tm_raster(palette = rev(viridisLite::magma(5)), breaks=c(0,2,20,200,2000,25000)) +
tm_shape(africountries) +
    tm_borders("white", lwd = .5) +
tm_shape(afrihighway) +
    tm_lines(col = "red") + 
tm_shape(africapitals) +
    tm_symbols(col = "blue", alpha=0.4, scale = .6 )+
tm_legend(show = FALSE)

```




